---
title: "instacart_training_dataset_prep"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Libraries
```{r setup, include=FALSE}
library(data.table)
library(dplyr)
library(lubridate)
```
#Mode Function
```{r}

getmode <- function(v){
  uv <- unique(v)
  uv[which.max(tabulate(match(v,uv)))]}

# Hours difference - hours run from 0 to 23
#circ_shift_dow<- function(x) { 
#     ifelse(x>3,x-4,x+3)}

hrDiff <- function(a,b) {
   diff = ifelse(a>b,a-b,b-a) 
   diff = min(diff,24-diff) 
   return(diff)}

dowDiff <- function(a,b) {
   diff = ifelse(a>b,a-b,b-a) 
   diff = min(diff,7-diff) 
   return(diff)}
```

# Read files
```{r}
orders =fread("data/orders.csv")
train_orders =fread("data/order_products__train.csv")
prior_orders=fread("data/order_products__prior.csv")
```

# Calculate number of days using the days_since_prev_order from each order
```{r}
orders %>% 
  mutate(days_since_prior_order = ifelse(is.na(days_since_prior_order),0,
                      days_since_prior_order))  %>% 
  mutate(O_day = ave(days_since_prior_order,user_id,FUN = cumsum)) %>% setDT()->orders
```

# Building user set;
```{r}
# train user set 
orders[orders$eval_set == "train",c(2,4,8)]  ->user_set


colnames(user_set)[2]  = "U_total_orders"
colnames(user_set)[3]  = "U_last_O_day"

```

# Extract all orders from user ; limit to last 100 orders if necessary
```{r}
user_set %>%
  head(10) ->user_set

user_set %>%
  left_join(orders[orders$eval_set=="prior",], by = "user_id") -> user_orders
user_orders[,-5] ->user_orders

#user_orders$order_dt = user_orders$start_dt + days(user_orders$O_day)
#user_orders[,names(user_orders) %in% c("user_id","order_id","order_number", #"order_dow","order_hour_of_day","days_since_prior_order","order_dt")] -> user_orders
```

# Build User_Features
         1. mostly ordered  hour of the day
         2. mostly ordered  dow
         3. average days between orders
```{r}
user_orders %>% 
  group_by(user_id) %>% 
     mutate(U_mostly_ord_hod = getmode(order_hour_of_day),
            U_mostly_ord_dow = getmode(order_dow)) %>%
  group_by(order_id) %>% 
  mutate(O_ord_dow_abs_var= dowDiff(order_dow,U_mostly_ord_dow),
         O_ord_hod_abs_var= hrDiff(order_hour_of_day,U_mostly_ord_hod) ) %>% 
  group_by(user_id,U_total_orders,U_mostly_ord_hod,U_mostly_ord_dow) %>% 
  summarise(U_mostly_ord_hod_var = median(O_ord_hod_abs_var),
            U_mostly_ord_dow_var = median(O_ord_dow_abs_var))->user_feature

#            U_dspo_mean = round(mean(days_since_prior_order)),
#            U_dspo_sd = round(sd(days_since_prior_order)),
#            U_days_since_last_purchase = as.numeric(U_last_order_day - O_day))
```

# Add order details from prior_order file
```{r}
products = fread("data/products.csv")

user_orders %>% 
  left_join(prior_orders, by="order_id") %>% setDT() -> user_prior_orders
rm(prior_orders)
rm(user_orders)

user_prior_orders %>% 
  left_join(products) -> user_prior_orders

# Limit the number of prior orders
user_prior_orders %>% 
#  filter(x-order_dt > 180) %>% 
  group_by(order_id) %>% 
  mutate(O_reord_count = sum(reordered),
        O_cart_size = n(),
        O_aisles = n_distinct(aisle_id),
        O_dept = n_distinct(department_id))  -> user_prior_orders

user_prior_orders %>% 
  group_by(order_id,O_cart_size,O_reord_count,O_aisles,O_dept) %>%
  mutate(O_reord_ratio_in_cart =  round(O_reord_count/O_cart_size,2) ,
         O_aisle_ratio_in_cart  = round(O_aisles / O_cart_size,2),
         O_dept_ratio_in_cart   = round(O_dept/O_cart_size,2)) -> orders_features
```

# More user features
```{r}
orders_features %>%
  group_by(user_id) %>% 
  summarise(U_avg_reord_ratio_in_cart = round(mean(O_reord_ratio_in_cart),2),
            U_avg_aise_ratio_in_cart  = round(mean(O_aisle_ratio_in_cart),2),
            U_avg_dept_ratio_in_cart  = round(mean(O_dept_ratio_in_cart),2),
            U_avg_cart_size       = round(mean(O_cart_size),1)) -> tmp_uf

user_feature %>% 
  merge(tmp_uf) -> user_feature
rm(tmp_uf)
rm(orders_features)
```

### Gather the complete list of products the customer ever purchased
```{r}
#user_prior_orders %>% 
#  group_by(user_id,product_id) %>% 
#  summarize(UP_last_purchased = max(O_day) ) ->user_prod_cross
#write.csv(user_prod_cross,"user_prod_cross_complete.csv")
```

# user_cross_product to identify first time orders
```{r}
user_prior_orders %>% group_by(user_id,product_id) ->user_prod_feature_for_new
```


# User_product_cross features
 Summarize at cross-product-level
      1, number of orders
      2. last day ordered
      3. mostly ordered dow 
      4. mostly ordered hod 
      5. repeat frequency
      6. percentage or orders of the product
      7. Days since last order placed as a ratio of reorder frequency
```{r}
user_prior_orders %>% 
  filter(U_last_O_day - O_day<=180) ->user_prior_orders

user_prior_orders %>% 
  group_by(user_id,product_id) %>% 
     mutate(UP_mostly_ord_dow = getmode(order_dow),
            UP_mostly_ord_hod = getmode(order_hour_of_day)) %>% 
  group_by(order_id,user_id,product_id)  %>% 
  mutate(UPO_ord_dow_abs_var = dowDiff(order_dow,UP_mostly_ord_dow),
           UPO_ord_hod_abs_var= hrDiff(order_hour_of_day,UP_mostly_ord_hod)) ->UP_tmp

UP_tmp %>% 
  group_by(user_id,product_id,#U_last_O_day,
           UP_mostly_ord_dow,UP_mostly_ord_hod,aisle_id,department_id) %>% 
  summarize(UP_prd_ord_count = n(),
            UP_days_since_last_purchase = first(U_last_O_day) - max(O_day),
          UP_ord_dow_abs_var = median(UPO_ord_dow_abs_var),
           UP_ord_hod_abs_var= median(UPO_ord_hod_abs_var),
            UP_pur_duration = max(O_day) - min(O_day),
            UP_add_to_cart_order =        
            round(mean(add_to_cart_order/O_cart_size),2)) %>% 
  mutate(UP_Pur_freq_days =round(UP_pur_duration/UP_prd_ord_count,1)) %>% 
  mutate(UP_pur_due = ifelse(UP_Pur_freq_days==0,0, 
              round(UP_days_since_last_purchase/UP_Pur_freq_days,2 ))) -> UP_tmp

 UP_tmp[,c("user_id"                    
 ,"product_id"                 
 ,"UP_prd_ord_count"           
 ,"UP_days_since_last_purchase"
 ,"UP_mostly_ord_dow"
 ,"UP_ord_dow_abs_var"         
 ,"UP_mostly_ord_hod"
 ,"UP_ord_hod_abs_var"         
 ,"UP_pur_duration"            
 ,"UP_add_to_cart_order"       
 ,"UP_Pur_freq_days"           
 ,"UP_pur_due"                 
 ,"aisle_id"                   
 ,"department_id")] ->user_prod_feature 


rm(UP_tmp)
```

# User_aisle features 
```{r}

user_prior_orders %>% 
  group_by(user_id,aisle_id) %>% 
     mutate(UA_mostly_ord_dow = getmode(order_dow),
            UA_mostly_ord_hod = getmode(order_hour_of_day)) %>% 
  group_by(order_id,user_id,aisle_id)  %>% 
  mutate(UAO_aisle_size = n(),
           UAO_ord_dow_abs_var = dowDiff(order_dow,UA_mostly_ord_dow),
           UAO_ord_hod_abs_var= hrDiff(order_hour_of_day,UA_mostly_ord_hod)) ->UA_tmp

UA_tmp %>% 
  group_by(user_id,aisle_id,#U_last_O_day,
           UA_mostly_ord_dow,UA_mostly_ord_hod) %>% 
  summarize(UA_aisle_ord_count = n(),
            UA_days_since_last_purchase = first(U_last_O_day) - max(O_day),
          UA_ord_dow_abs_var = median(UAO_ord_dow_abs_var),
           UA_ord_hod_abs_var= median(UAO_ord_hod_abs_var),
            UA_pur_duration = max(O_day) - min(O_day),
            UA_add_to_cart_order = round(mean(add_to_cart_order/O_cart_size),2)) %>% 
  mutate(UA_Pur_freq_days =round(UA_pur_duration/UA_aisle_ord_count,1)) %>% 
  mutate(UA_pur_due = ifelse(UA_Pur_freq_days==0,0, 
              round(UA_days_since_last_purchase/UA_Pur_freq_days,2 ))) -> UA_tmp

 UA_tmp[,c("user_id"                    
 ,"aisle_id"                   
 ,"UA_aisle_ord_count"           
 ,"UA_days_since_last_purchase"
 ,"UA_mostly_ord_dow"
 ,"UA_ord_dow_abs_var"         
 ,"UA_mostly_ord_hod"
 ,"UA_ord_hod_abs_var"         
 ,"UA_pur_duration"            
 ,"UA_add_to_cart_order"       
 ,"UA_Pur_freq_days"           
 ,"UA_pur_due")] ->user_aisle_feature 

rm(UA_tmp)
```

# User Department features 
```{r}

user_prior_orders %>% 
  group_by(user_id,department_id) %>% 
     mutate(UD_mostly_ord_dow = getmode(order_dow),
            UD_mostly_ord_hod = getmode(order_hour_of_day)) %>% 
  group_by(order_id,user_id,department_id)  %>% 
  mutate(UDO_dept_size = n(),
           UDO_ord_dow_abs_var = dowDiff(order_dow,UD_mostly_ord_dow),
           UDO_ord_hod_abs_var= hrDiff(order_hour_of_day,UD_mostly_ord_hod)) ->UD_tmp

UD_tmp %>% 
  group_by(user_id,department_id,#U_last_O_day,
           UD_mostly_ord_dow,UD_mostly_ord_hod) %>% 
  summarize(UD_dept_ord_count = n(),
            UD_days_since_last_purchase = first(U_last_O_day) - max(O_day),
          UD_ord_dow_abs_var = median(UDO_ord_dow_abs_var),
           UD_ord_hod_abs_var= median(UDO_ord_hod_abs_var),
            UD_pur_duration = max(O_day) - min(O_day),
            UD_add_to_cart_order = round(mean(add_to_cart_order/O_cart_size),2)) %>% 
  mutate(UD_Pur_freq_days =round(UD_pur_duration/UD_dept_ord_count,1)) %>% 
  mutate(UD_pur_due = ifelse(UD_Pur_freq_days==0,0, 
              round(UD_days_since_last_purchase/UD_Pur_freq_days,2 ))) -> UD_tmp

 UD_tmp[,c("user_id"                    
 ,"department_id"                   
 ,"UD_dept_ord_count"           
 ,"UD_days_since_last_purchase"
 ,"UD_mostly_ord_dow"
 ,"UD_ord_dow_abs_var"         
 ,"UD_mostly_ord_hod"
 ,"UD_ord_hod_abs_var"         
 ,"UD_pur_duration"            
 ,"UD_add_to_cart_order"       
 ,"UD_Pur_freq_days"           
 ,"UD_pur_due")] ->user_dept_feature 

rm(UD_tmp)
rm(user_prior_orders)
```

# Build Future order_features
```{r}
orders[orders$eval_set == "train",c(2,5,6,7,8)]  ->user_FUT_features

colnames(user_FUT_features)=c("user_id","FUT_order_dow","FUT_order_hod","FUT_days_since","FUT_O_day")

#write.csv(user_FUT_features, "user_FUT_features.csv",col.names = NA)

```

# Training dataset 
```{r}
user_prod_feature %>% 
  left_join(user_feature, by="user_id") %>% 
  left_join(user_aisle_feature, by = c("user_id", "aisle_id")) %>% 
  left_join(user_dept_feature, by = c("user_id", "department_id"))  %>% 
  left_join(user_FUT_features,by=c("user_id"))  %>% setDT() ->prior_orders_features

rm(user_prod_feature)
```

# Prepare future orders(outcomes) for training
```{r}
# train user set 
orders[orders$eval_set == "train",c(1,2)]  ->future_ord
future_ord[future_ord$user_id %in% user_set$user_id, ]  ->future_ord

future_ord %>% 
  left_join(train_orders) %>% 
  select(user_id,product_id,reordered) ->future_ord

```


#Final training file for predicting reorders
```{r}
prior_orders_features %>%                                 
  left_join(future_ord,by = c("user_id", "product_id")) %>% 
  mutate(ORDERED =  ifelse(is.na(reordered),0,1) ) -> training_data


apply(training_data[,c("UP_mostly_ord_hod","FUT_order_hod")], 1 ,
      function(x) hrDiff(x[1],x[2])) ->
  training_data$UP_hod_diff

apply(training_data[,c("UP_mostly_ord_dow","FUT_order_dow")], 1 ,
      function(x) dowDiff(x[1],x[2])) -> 
  training_data$UP_dow_diff

write.csv(training_data, "training_data10k.csv")
```

# Training data for New orders
```{r}
future_ord %>% 
  left_join(products,by="product_id") %>% 
  left_join(user_feature, by="user_id") %>% 
  left_join(user_aisle_feature, by = c("user_id", "aisle_id")) %>% 
  left_join(user_dept_feature, by = c("user_id", "department_id"))  %>% 
  left_join(user_FUT_features,by=c("user_id"))  %>% setDT()  %>% 
  anti_join(user_prod_feature_for_new, by=c("user_id","product_id")) ->new_order_training_data

rm(user_feature)
rm(user_aisle_feature)
rm(user_dept_feature)
rm(user_FUT_features)


write.csv(new_order_training_data,"new_order_training_data.csv")
```
